---
title: "Tata steel analysis"
output: html_document
---
In this project, I aim to implement some concepts of econophysics, later updates would include adding in Black-Scholes model and making a portfolio optimization model

```{r}
pacman::p_load(pacman,dplyr,GGally, ggplot2, ggthemes, ggvis, httr, lubridate, plotly, rio, rmarkdown, shiny, stringr, tidyr, fitdistrplus)

library(datasets)
library(tidyquant)
library(tseries)
library(zoo)
```

We analyse the price of Tata Steel for the last 5 years and try to predict the future trend.

```{r}
getSymbols("TATASTEEL.BO", from = "2015-01-01", to= "2020-12-31",auto.assign = TRUE,warnings = FALSE)

tatasteel_price <- (TATASTEEL.BO$TATASTEEL.BO.Adjusted)

tatasteel_price <- as.numeric(na.omit(tatasteel_price))
plot(ts(tatasteel_price))
```

First we check for correlation between the stock values in order to know if there is some kind of meaningful data that we can use for analysis.
For that we use the auto correlation function

```{r}
acf(tatasteel_price, plot = T)
```

we see that there is a significant correlation even beyond 30 lag. But we don't find the "true" 
correlation between a current time spot and the time spot just before it as they are influenced by earlier time spots. 
To find the correlation between these two, we use the PACF plot.

```{r}
pacf(tatasteel_price, plot = T)
```

we find significant correlation at lag 1 and then it falls below the 95% confidence line indicating an auto regressive term with lag 1 but a correlation of 1 can also signify a random walk, hence we check the first difference of the data

```{r}
diff_return <- as.matrix(diff(tatasteel_price))

plot(ts(diff_return))

acf(diff_return)

pacf(diff_return)

Box.test(diff_return, lag= 20, type = c("Ljung-Box"))
```

We see that there is a significant correlation at lag 7 for both acf and pacf
Then we ran the Ljung box test to check if the correlations are not due to chance. as the p-value<0.05 we can reject the null hypothesis and assume that our data is in fact, correlated.



RANDOM WALK THEORY : "THERE IS NO METHOD FOR PREDICTION OF STOCK PRICES AS THERE IS TOO MUCH OF RANDOM MOVEMENT,INVESTORS ARE BETTER OFF INVESTING IN STOCKS WHICH HAVE GENERAL TREND OF MARKET."

To test this theory we generate a random walk with fixed amount of movement


```{r}
rnd_price <- matrix(NA, nrow = length(tatasteel_price), ncol = 1)

rnd_price[1,] = 650

for(i in 2: length(tatasteel_price)){
  toss <- sample(c("H","T"),1,replace = TRUE,prob =c(0.5,0.5))
  if(toss == "H") rnd_price[i,]= rnd_price[i-1,]+ 100
  if(toss == "T") rnd_price[i,]= rnd_price[i-1,]-100
}

par(mfrow=c(2,1))
par("mar")

plot(ts(tatasteel_price))
plot(ts(rnd_price))
```


Looking at the plots, we see that random walk is not able to show a significant trend of the original price so instead of using random steps for price movement, we can use return 


Using log return:


```{r}
set.seed(101)
log_return_tatasteel <- diff(log(tatasteel_price))
log_return_tatasteel <- as.numeric(na.omit(log_return_tatasteel))
logis <- fitdist(log_return_tatasteel,'logis',method = "mle")
plot(ts(log_return_tatasteel))
plot(density(log_return_tatasteel))
acf(log_return_tatasteel)
plot(logis)
```
Hence we see that logistic distribution is a good fit to our data.

Geometric Random Walk:
P[T]=P[0]exp(r(t)+...+r(1))
```{r}
random_log_return <- cumsum(rlogis(length(tatasteel_price),0.000438,0.013333)) # This will calculate the sum of data point
log_price <- rep(NA,755)
log_price[1] <- 260
for(i in 2:length(tatasteel_price)){
  log_price[i] <- log_price[1] * exp(random_log_return[i])
}
plot(ts(tatasteel_price))
plot(ts(log_price),col = 'red')
```

We see that the log return is able to grasp a bit of the market trend. 
Now we need to check if the distribution is going to stay the same as time goes on, we can do that if we can prove that our log return is a stationary time series.

Stationary time series:
mean and variance does not grow without bound
covariance depends upon time

```{r}
MEAN_CORRELATION_VAR <- function(data,lag,sample_size){
  if(length(data) %% sample_size == 0){
    l <- length(data)/sample_size
    mean_value <- rep(NA,l)
    acf_value <- rep(NA,l)
    var_value <- var(NA,l)
    iteration <- 1
    step_size <- 1
    while(iteration <l){
      sample_data <- data[step_size:(step_size+sample_size)] #data is broken into lengths of sample size
      mean_value[iteration] <- mean(sample_data) # mean of sample
      acf_mu <- acf(sample_data,lag.max = lag,plot = FALSE) #covariance of sample
      acf_value[iteration] <- acf_mu$acf[lag+1]
      var_value[iteration] <- var(sample_data) #VARIANCE OF SAMPLE
      step_size <- step_size+4
      iteration <- iteration+1
    }
      
  }
  else
    print("Error")
  
 
  complie_list <- cbind(mean_value,acf_value,var_value)
}
#Function to check if time series is stationary or not
```

Conditions to be satisfied to apply geometric random walk:
1.price should not be stationary
2.return should be stationary
3.correlation should be there among return data points

```{r}
set.seed(101)
sample_price <- tatasteel_price[1:750] 
tatasteel_stationary <- MEAN_CORRELATION_VAR(sample_price,5,50)
tatasteel_stationary <- na.omit(tatasteel_stationary)
plot(ts(tatasteel_stationary[,1]))
title(main = "MEAN_MOVEMENT")
plot(ts(tatasteel_stationary[,2]))
abline(h=0,lwd=2,col='red')
title(main = "CORRELATION_MOVEMENT")
plot(ts(tatasteel_stationary[,3]))
title(main = "VARIANCE_MOVEMENT")
```
As the variance and mean is non constant, we can say that the series is not stationary

Now checking if log return is stationary:

```{r}
set.seed(101)
mod_log_return <- log_return_tatasteel[1:750]
return_analysis <- MEAN_CORRELATION_VAR(mod_log_return,5,50)
plot(ts(return_analysis[,1]))
abline(h=0,lwd=2,col='red')
title(main = "MEAN_MOVEMENT")
plot(ts(return_analysis[,2]))
abline(h=-0.2,lwd=2,col='red')
title(main = "CORRELATION_MOVEMENT")
plot(ts(return_analysis[,3]))
abline(h=0.0006,lwd=2,col='red')
title("VARIANCE_MOVEMENT")
```
Here we can see that the variance and mean jump up first but then come back to a low value hence as they are not growing exponentially we can say that our log returns are stationary. However, there's a need of robust testing to verify this.
Here, augmented Dickey-Fuller Test would be useful to determine if the series is stationary or not.
It has two hypothesis:
$H_{o}$: The time series is unit root time series (non stationary series)
$H_{1}$: The series is stationary

Greater p-value implies towards $H_{o}$ hypothesis

```{r}
library(tseries)
adf.test(log(tatasteel_price))
adf.test(log_return_tatasteel)
```
Hence we can see that observations about both series is correct.
We also need to rule out the possibility of our log return series being noise

```{r}
Box.test(log_return_tatasteel,lag = 20,type = "Ljung")
```
We see that there is correlation between the data so we can apply geometric random motion for prediction of future prices

Geometric Brownian Motion:

$\hat{\mu} = {(\overline{r}+s^2/2)}/{\Delta t}$

$\hat{\sigma} = s/{\Delta t}$


```{r}
n <- length(tatasteel_price)
sim.size = 500
rbar<-mean(log_return_tatasteel)
s<-sd(log_return_tatasteel)
delta_t<-1
mu_hat<-rbar+s^2/2
set.seed(321)
## Simulate log-return from Normal distribution
rt.sim<-rnorm(sim.size,mean=(mu_hat-s^2/2),sd=s)
tatasteel.sim <-rep(NA,sim.size)
tatasteel.sim[1] <- tatasteel_price[n]*exp(rt.sim[1])

for(i in 2: sim.size) 
  tatasteel.sim[i] <- tatasteel.sim[i-1]*exp(rt.sim[i])


yl<-min(tatasteel_price)*0.85
yu<-max(tatasteel_price)*1.9
plot(ts(tatasteel_price),xlim=c(0,(n+sim.size)),ylim=c(yl,yu))
lines((n+1):(n+sim.size),tatasteel.sim,col="red",lwd=2)
grid(col="black",lwd=2)
getSymbols("TATASTEEL.BO", from = "2020-01-01", to= "2021-8-1",auto.assign = TRUE,warnings = FALSE)

tatasteel_price_new <- (TATASTEEL.BO$TATASTEEL.BO.Adjusted)

tatasteel_price_new <- as.numeric(na.omit(tatasteel_price_new))

lines((n+1):(n+length(tatasteel_price_new)),tatasteel_price_new,col="green",lwd=2)
grid(col="black",lwd=2)

```
For verification, you can actually see the stock market graph for tata steel for the year 2021 which is the green line and the prediction is the red line






Unload all packages
```{r}

p_unload(all)

detach("package:datasets",unload = TRUE)
```

